### 2015-06-05 mdp.tex
modified mdp.tex

%!TEX root =  autocontgrlp.tex
\section{Markov Decision Processes (MDPs)}
In this section, we briefly discuss the basics of Markov Decision Processes (MDPs) (the reader is referred to \cite{BertB,Puter} for a detailed treatment).\\
An MDP is a $4$-tuple $\langle S,A,P,g\rangle$, where $S$ is the state space, $A$ is the action space, $P$ is the probability transition kernel and $g$ is the reward function. We consider MDPs with large but finite number of states. We also assume that the number of actions is finite.
Without the loss of generality, we set  $S=\{1,2,\ldots,n\}$  and the action set is given by $A=\{1,2,\ldots,d\}$. For simplicity, we assume that all actions are feasible in all states. The probability transition kernel $P$ specifies the probability $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$. We denote the reward obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.

A policy $\mu$ specifies the action selection mechanism, and is described by the sequence $\mu=(u_1,u_2,\ldots,u_n,\ldots)$, where $u_n\colon S \ra A$ for all $n \geq 0$. \todoc{This is somewhat limited; normally we should allow history dependent,
and also stochastic policies.} 
A stationary deterministic policy (SDP) is one where $u_n\equiv u$ for all $n\geq 0$ for some $u\colon S \ra A$. By abuse of notation we denote the SDP by $u$ itself instead of $\mu$. In the setting that we consider, one can find an SDP that is optimal \cite{BertB,Puter}. In this paper, we restrict our focus to the class $U$ of SDPs.  Under an SDP $u$, the MDP is a Markov chain with probability transition kernel $P_u$.

Given an SDP $u$, the infinite horizon discounted reward corresponding to state $s$ under $u$ is denoted by $J_u(s)$ and is defined by

@@

J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n), n= 0,1,\dots],\nn

@@

where $\alpha \in (0,1)$ is a given discount factor. Here $J_u(s)$ is known as the value of the state $s$ under the SDP $u$, and the vector quantity $J_u\stackrel{\Delta}{=}(J_u(s))_{s\in S}\in R^n$ is called the value-function corresponding to the SDP $u$.\\
The optimal policy $u^*$ is any policy whose value is the largest possible under any policy: 
$J_{u^*}(s) = \max_{u\in U} J_u(s)$ for all $s\in S$. Such $u^*$ exists and is well defined in the case of infinite horizon discounted reward MDP \cite{Puter}.

@@

 The optimal value-function $J^*$ is the one obtained under the optimal policy, i.e., $J^*=J_{u^*}$. \todoc{Probably better to define first $J^*$ as the pointwise max and then the optimal policy's definition could use $J^*$.}
Given an MDP, our aim is to find the optimal value function $J^*$ and the optimal policy $u^*$. The optimal policy and value function obey the Bellman equation (BE): for all $ s \in S$, \todoc{Non-unique maximizers in the policy def?}

@@

@@

Note that Lemmas~\ref{maxnorm}-\ref{shift} also hold for the $Q$-Bellman operator $H$.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Markov Decision Processes (MDPs)}
In this section, we briefly discuss the basics of Markov Decision Processes (MDPs) (the reader is referred to \cite{BertB,Puter} for a detailed treatment).\\
An MDP is a $4$-tuple $\langle S,A,P,g\rangle$, where $S$ is the state space, $A$ is the action space, $P$ is the probability transition kernel and $g$ is the reward function. We consider MDPs with large but finite number of states. We also assume that the number of actions is finite.
Without the loss of generality, we set  $S=\{1,2,\ldots,n\}$  and the action set is given by $A=\{1,2,\ldots,d\}$. For simplicity, we assume that all actions are feasible in all states. The probability transition kernel $P$ specifies the probability $p_a(s,s')$ of transitioning from state $s$ to state $s'$ under the action $a$. We denote the reward obtained for performing action $a\in A$ in state $s\in S$ by $g_a(s)$.
\todoc{This is somewhat limited; normally we should allow history dependent,
and also stochastic policies.} 
By a policy, we mean a sequence $\pi=\{\pi_0,\ldots,\pi_n,\ldots\}$ of functions $\pi_i, i\geq 0$ that describe the manner in which an action is picked in a given state at time $i$. The most general class of policies is the class of \emph{history-dependent randomized policies} denoted by $\pi$ and is defined below.
Let $h_n=\{s_0,\ldots,s_n, \pi_0, \ldots, \pi_{n-1}\}$ denote the \emph{history}. A policy $\pi$ is said to be a \emph{history-dependent randomized} policy if $\pi_n,n\geq 0$ is such that for every $s\in S$, $\pi_n(s,\cdot|h_n)$ is a probability distribution over the set of actions.
We denote the class of policies by $\Pi$. The infinite horizon discounted reward corresponding to state $s$ under a policy $\pi$ is denoted by $J_\pi(s)$ and is defined by

@@

J_\pi(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|\pi],\nn

@@

where $a_{n} \sim \pi_n(s,\cdot)$ and $\alpha \in (0,1)$ is a given discount factor.  Here $J_\pi(s)$ is known as the value of the state $s$ under the policy $\pi$, and the vector quantity $J_\pi\stackrel{\Delta}{=}(J_\pi(s))_{s\in S}\in R^n$ is called the value-function corresponding to the policy $\pi$. The optimal value function $J^*$ is defined as $J^*(s)\stackrel{def}{=}\underset{\pi \in \Pi}{\max} J_\pi(s)$.\par
A stationary deterministic policy (SDP) is one where $\pi_n\equiv u$ for all $n\geq 0$ for some $u\colon S \ra A$. By abuse of notation we denote the SDP by $u$ itself instead of $\pi$. In the setting that we consider, one can find an SDP that is optimal \cite{BertB,Puter}, i.e., there exists an SDP $u^*$ such that $J_{u^*}(s)=J^*(s),\forall s\in S$. In this paper, we restrict our focus to the class $U$ of SDPs and without loss of generality we assume that there exists a unique SDP $u^*$ that is optimal. Under a stationary policy $u$ (or $\pi$), the MDP is a Markov chain and we denote its probability transition kernel by $P_u=(p_{u(i)}(i,j),i,j=1,\ldots,n)$ (or $P_\pi=(p_{\pi(i)}(i,j),i,j=1,\ldots,n)$, where $p_{\pi(i)}(i,j)=\sum_{a\in A}\pi(i,a)p_a(i,j)$ and $\pi(i)=(\pi(i,a), a\in A)$).\par

@@

%The optimal value-function $J^*$ is the one obtained under the optimal policy, i.e., $J^*=J_{u^*}$. 
\todoc{Probably better to define first $J^*$ as the pointwise max and then the optimal policy's definition could use $J^*$.}
Given an MDP, our aim is to find the optimal value function $J^*$ and the optimal policy SDP $u^*$. The optimal policy and value function obey the Bellman equation (BE): for all $ s \in S$, \todoc{Non-unique maximizers in the policy def?}

@@

\begin{definition}\label{bellactval}

@@

Note that Lemmas~\ref{maxnorm}-\ref{shift} also hold for the Bellman operator $H$ defined for the action values in Definition~\ref{bellactval}.\par
